---
title: "Figure 2; Inhibition of HSP90 exposes cryptic variation in traits diverging between populations in an invasive ant"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: true
    number_sections: true
    df_print: paged
  pdf_document: default
---
<STYLE TYPE="text/css">
<!--
  td{
    font-family: Arial; 
    font-size: 8pt;
    padding:0px;
    cellpadding="0";
    cellspacing="0"
  }
  th {
    font-family: Arial; 
    font-size: 8pt;
    height: 20px;
    font-weight: bold;
    text-align: right;
    background-color: #ccccff;
  }
  table { 
    border-spacing: 0px;
    border-collapse: collapse;
  }
--->
</STYLE>
# Setup environment

## install libraries
```{r}
# install.packages("devtools")
library(devtools)
install_github("thomasp85/patchwork")
```

## load libraries
```{r}
library(ggplot2)
source("~/sciebo/librarySchrader.R")
library(cowplot)
library(rgr)
library("corrplot")
library(factoextra)
library(FactoMineR)
library(ggExtra)
library(sjstats)
library("gridExtra")
library(patchwork)
library(cvequality)
library(coin)
```

# Process data
## Load morphometric data
```{r}
data<-read.csv("/Users/lukas/sciebo/Projects/BSc/Miles Winter/paper/paper/lukas/Figures/Fig2/morphometry.tsv",sep="\t",stringsAsFactors = F)
data$group <- factor(data$group, levels = c("C","T","BR","TN"))
summary(data$group)

```

## Define colors
```{r}
data$group<-as.factor(data$group)
data$col<-NA
data$col[data$group=="C"]<-goodCols[1]
data$col[data$group=="T"]<-goodCols[3]
data$col[data$group=="TN"]<-goodCols[5]
data$col[data$group=="BR"]<-goodCols[7]

```
## Define experiments
```{r}
data$exp<-as.factor(ifelse(data$group=="BR"|data$group=="TN","nat. pop.","17-DMAG"))
```

## Calculate simple correlations
```{r}
gg1 <- ggplot(data, aes(x = ML, y = SP,colour = group)) + geom_point(alpha=1,cex=2,color=data$col) 
gg1<-gg1 + geom_density_2d(alpha=.2,cex=.3) +  scale_color_manual(values = c(goodCols[c(1,3,5,7)]))+theme_light() +   geom_smooth(method='lm',alpha=.2)

gg8 <- ggplot(data, aes(x = EY, y = HL,colour = group)) + geom_point(alpha=1,cex=2,color=data$col) 
gg8 <-gg8 + geom_density_2d(alpha=.2,cex=.3) +  scale_color_manual(values = c(goodCols[c(1,3,5,7)]))+theme_light() +   geom_smooth(method='lm',alpha=.2,cex=.8)

gg10 <- ggplot(data, aes(x = HL, y = CL,colour = group)) + geom_point(alpha=1,cex=2,color=data$col) 
gg10 <- gg10 + geom_density_2d(alpha=.2,cex=.3) +  scale_color_manual(values = c(goodCols[c(1,3,5,7)]))+theme_light() +   geom_smooth(method='lm',alpha=.2,cex=.8)

```
## Plot simple correlations
```{r warning=FALSE}
gg1
gg8
gg10
```


## Create boxplots for all 4 groups
```{r}
alphaVal<-0.5
dotS<-0.5
axtexsize<-10
p1 <- ggplot(data, aes(x=group, y=ML)) + 
  geom_boxplot(outlier.shape = NA) + facet_grid(. ~ exp,scales = "free",switch="x") + theme(strip.background = element_rect(fill=rgb(0,0,0,0)),strip.text.x=element_text(size=7,vjust=.9))
p1<-p1 + geom_dotplot(binaxis='y', stackdir='center', dotsize=dotS, aes(x=group, y=ML,fill=group,color=group)) +
  theme(legend.position="none") +
  scale_fill_manual(values=addalpha(c(rep("black",4)),alphaVal))+
  scale_color_manual(values=rep(NA,4)) +
  theme(axis.title.x=element_blank(),axis.text=element_text(size=axtexsize),axis.title.y = element_text(margin = margin(t = 0, r = -15, b = 0, l = 0)))  + 
  scale_y_continuous(breaks = c(390, 500))

p2 <- ggplot(data, aes(x=group, y=SP)) + 
  geom_boxplot(outlier.shape = NA) + facet_grid(. ~ exp,scales = "free",switch="x") + theme(strip.background = element_rect(fill=rgb(0,0,0,0)),strip.text.x=element_text(size=7,vjust=.9))
p2<-p2 + geom_dotplot(binaxis='y', stackdir='center', dotsize=dotS, aes(x=group, y=SP,fill=group,color=group)) + 
  theme(legend.position="none") + 
  scale_fill_manual(values=addalpha(c(rep("black",4)),alphaVal)) +
  scale_color_manual(values=rep(NA,4)) +
  theme(axis.title.x=element_blank(),axis.text=element_text(size=axtexsize),axis.title.y = element_text(margin = margin(t = 0, r = -15, b = 0, l = 0)))  + 
  scale_y_continuous(breaks = c(100, 165))+ expand_limits(y=170)

p3 <- ggplot(data, aes(x=group, y=EY)) + 
  geom_boxplot(outlier.shape = NA) + facet_grid(. ~ exp,scales = "free",switch="x") + theme(strip.background = element_rect(fill=rgb(0,0,0,0)),strip.text.x=element_text(size=7,vjust=.9))

p3 <- p3 + geom_dotplot(binaxis='y', stackdir='center', dotsize=dotS, aes(x=group, y=EY,fill=group,color=group)) + 
  theme(legend.position="none") +
  scale_fill_manual(values=addalpha(c(rep("black",4)),alphaVal))+ 
  scale_color_manual(values=rep(NA,4)) +
  theme(axis.title.x=element_blank(),axis.text=element_text(size=axtexsize),axis.title.y = element_text(margin = margin(t = 0, r = -15, b = 0, l = 0)))  + 
  scale_y_continuous(breaks = c(370, 420)) + expand_limits(y=430)
  

p4 <- ggplot(data, aes(x=group, y=HL)) + 
  geom_boxplot(outlier.shape = NA) + facet_grid(. ~ exp,scales = "free",switch="x") + theme(strip.background = element_rect(fill=rgb(0,0,0,0)),strip.text.x=element_text(size=7,vjust=.9))
p4<-p4 + geom_dotplot(binaxis='y', stackdir='center', dotsize=dotS, aes(x=group, y=HL,fill=group,color=group)) + 
  theme(legend.position="none") + 
  scale_fill_manual(values=addalpha(c(rep("black",4)),alphaVal))+ 
  scale_color_manual(values=rep(NA,4)) +
  theme(axis.title.x=element_blank(),axis.text=element_text(size=axtexsize),axis.title.y = element_text(margin = margin(t = 0, r = -15, b = 0, l = 0)))  + 
  scale_y_continuous(breaks = c(375, 475)) + expand_limits(y=480)

p5 <- ggplot(data, aes(x=group, y=CL)) + 
  geom_boxplot(outlier.shape = NA) + facet_grid(. ~ exp,scales = "free",switch="x") + theme(strip.background = element_rect(fill=rgb(0,0,0,0)),strip.text.x=element_text(size=7,vjust=.9))
p5l<-p5 + geom_dotplot(binaxis='y', stackdir='center', dotsize=dotS, aes(x=group, y=CL,fill=group,color=group)) + 
  guides(fill=guide_legend(ncol=1)) + 
  scale_fill_manual(values=addalpha(c(rep("black",4)),alphaVal))+
  scale_color_manual(values=rep(NA,4)) + 
  theme(axis.title.x=element_blank(),axis.text=element_text(size=axtexsize),axis.title.y = element_text(margin = margin(t = 0, r = -15, b = 0, l = 0)))  + 
  scale_y_continuous(breaks = c(95, 120))

p5 <- p5l+ theme(legend.position="none")
legend <- get_legend(p5l)
```

## Calculate F-test for Treatment vs Control
```{r}
# f-test
dat2<-split(data,f=data$group)
tests<-list()
tests[["ML"]]<-var.test(dat2[["T"]]$ML,dat2[["C"]]$ML)
tests[["SP"]]<-var.test(dat2[["T"]]$SP,dat2[["C"]]$SP)
tests[["EY"]]<-var.test(dat2[["T"]]$EY,dat2[["C"]]$EY)
tests[["HL"]]<-var.test(dat2[["T"]]$HL,dat2[["C"]]$HL)
tests[["CL"]]<-var.test(dat2[["T"]]$CL,dat2[["C"]]$CL)
```

## Calculate mslr Tests for Treatment vs Control
```{r}
#Marwick and Krishnamoorthy 2016 
mslrTests<-list()

tmp<-subset(data,!is.na(ML) & (group=="T"|group=="C"))
tmp$group<-as.factor(as.character(tmp$group))
mslrTests[["ML"]]<-mslr_test(nr=10000,tmp$ML,tmp$group)

tmp<-subset(data,!is.na(SP)& (group == "T"|group == "C"))
tmp$group<-as.factor(as.character(tmp$group))
mslrTests[["SP"]]<-mslr_test(nr=10000,tmp$SP,tmp$group)

tmp<-subset(data,!is.na(CL)& (group == "T"|group == "C"))
tmp$group<-as.factor(as.character(tmp$group))
mslrTests[["CL"]]<-mslr_test(nr=10000,tmp$CL,tmp$group)

tmp<-subset(data,!is.na(EY)& (group == "T"|group == "C"))
tmp$group<-as.factor(as.character(tmp$group))
mslrTests[["EY"]]<-mslr_test(nr=10000,tmp$EY,tmp$group)

tmp<-subset(data,!is.na(HL)& (group == "T"|group == "C"))
tmp$group<-as.factor(as.character(tmp$group))
mslrTests[["HL"]]<-mslr_test(nr=10000,tmp$HL,tmp$group)

```

## Generate Table object with all relevant information
```{r}
means<-list()
sds<-list()
medians<-list()
mins<-list()
maxs<-list()
cvs<-list()
vars<-list()

for (p in c("T","C")){
  for (q in c('ML','SP','EY','HL','CL')){
    means[[p]][q]<-mean(dat2[[p]][,q],na.rm=T)
    sds[[p]][q]<-sd(dat2[[p]][,q],na.rm=T)
    medians[[p]][q]<-median(dat2[[p]][,q],na.rm=T)
    mins[[p]][q]<-min(dat2[[p]][,q],na.rm=T)
    maxs[[p]][q]<-max(dat2[[p]][,q],na.rm=T)
    cvs[[p]][q]<-cv(dat2[[p]][,q])
    vars[[p]][q]<-var(dat2[[p]][,q],na.rm=T)
  }
}
m1<-do.call(cbind,means)
colnames(m1)<-c("T_mean","C_mean")
m2<-do.call(cbind,sds)
colnames(m2)<-c("T_sd","C_sd")
m3<-do.call(cbind,medians)
colnames(m3)<-c("T_median","C_median")
m4<-do.call(cbind,mins)
colnames(m4)<-c("T_min","C_min")
m5<-do.call(cbind,maxs)
colnames(m5)<-c("T_max","C_max")
m6<-do.call(cbind,cvs)
colnames(m6)<-c("T_cv","C_cv")
m7<-do.call(cbind,vars)
colnames(m7)<-c("T_var","C_var")
all<-cbind(m1,m2,m3,m4,m5,m6,m7)

fstats<-vector()
pvalue<-vector()
mslrStats<-vector()
mslrP<-vector()
for (q in c('ML','SP','EY','HL','CL')){
  fstats[q]<-(tests[[q]]$statistic)
  pvalue[q]<-(tests[[q]]$p.value)
  mslrStats[q]<-mslrTests[[q]]$MSLRT
  mslrP[q]<-mslrTests[[q]]$p_value
}
fstats<-data.frame(fstats)
pvalue<-data.frame(pvalue)
mslrStats<-data.frame(mslrStats)
mslrP<-data.frame(mslrP)
all<-merge(all,mslrStats,by="row.names")
row.names(all)<-all$Row.names
all2<-merge(all[,-1],mslrP,by="row.names")
colnames(all2)[1]<-"measurement"
write.table(all2,"/Users/lukas/sciebo/Projects/BSc/Miles Winter/paper/paper/lukas/Figures/Fig2/morphometryTable.experiment.tsv",sep="\t",quote=F,row.names = F)

all2
```



## Calculate deviation from the mean
```{r}
data$ML_d<-NA
data$CL_d<-NA
data$EY_d<-NA
data$SP_d<-NA
data$HL_d<-NA

data$ML_d[data$group=="C"]<-abs(data$ML[data$group=="C"]-mean(data$ML[data$group=="C"],na.rm=TRUE))
data$CL_d[data$group=="C"]<-abs(data$CL[data$group=="C"]-mean(data$CL[data$group=="C"],na.rm=T))
data$EY_d[data$group=="C"]<-abs(data$EY[data$group=="C"]-mean(data$EY[data$group=="C"],na.rm=T))
data$SP_d[data$group=="C"]<-abs(data$SP[data$group=="C"]-mean(data$SP[data$group=="C"],na.rm=T))
data$HL_d[data$group=="C"]<-abs(data$HL[data$group=="C"]-mean(data$HL[data$group=="C"],na.rm=T))

data$ML_d[data$group=="T"]<-abs(data$ML[data$group=="T"]-mean(data$ML[data$group=="T"],na.rm=T))
data$CL_d[data$group=="T"]<-abs(data$CL[data$group=="T"]-mean(data$CL[data$group=="T"],na.rm=T))
data$EY_d[data$group=="T"]<-abs(data$EY[data$group=="T"]-mean(data$EY[data$group=="T"],na.rm=T))
data$SP_d[data$group=="T"]<-abs(data$SP[data$group=="T"]-mean(data$SP[data$group=="T"],na.rm=T))
data$HL_d[data$group=="T"]<-abs(data$HL[data$group=="T"]-mean(data$HL[data$group=="T"],na.rm=T))

data$ML_d[data$group=="BR"]<-abs(data$ML[data$group=="BR"]-mean(data$ML[data$group=="BR"],na.rm=T))
data$CL_d[data$group=="BR"]<-abs(data$CL[data$group=="BR"]-mean(data$CL[data$group=="BR"],na.rm=T))
data$EY_d[data$group=="BR"]<-abs(data$EY[data$group=="BR"]-mean(data$EY[data$group=="BR"],na.rm=T))
data$SP_d[data$group=="BR"]<-abs(data$SP[data$group=="BR"]-mean(data$SP[data$group=="BR"],na.rm=T))
data$HL_d[data$group=="BR"]<-abs(data$HL[data$group=="BR"]-mean(data$HL[data$group=="BR"],na.rm=T))

data$ML_d[data$group=="TN"]<-abs(data$ML[data$group=="TN"]-mean(data$ML[data$group=="TN"],na.rm=T))
data$CL_d[data$group=="TN"]<-abs(data$CL[data$group=="TN"]-mean(data$CL[data$group=="TN"],na.rm=T))
data$EY_d[data$group=="TN"]<-abs(data$EY[data$group=="TN"]-mean(data$EY[data$group=="TN"],na.rm=T))
data$SP_d[data$group=="TN"]<-abs(data$SP[data$group=="TN"]-mean(data$SP[data$group=="TN"],na.rm=T))
data$HL_d[data$group=="TN"]<-abs(data$HL[data$group=="TN"]-mean(data$HL[data$group=="TN"],na.rm=T))

colnames(data)[colnames(data)=="HL_d"]<-"dHL"
colnames(data)[colnames(data)=="CL_d"]<-"dCL"
colnames(data)[colnames(data)=="SP_d"]<-"dSP"
colnames(data)[colnames(data)=="ML_d"]<-"dML"
colnames(data)[colnames(data)=="EY_d"]<-"dEY"
```

# Focus on population comparison
## Create subset for only population specific data
```{r}
dataPop<-subset(data,group=="TN"|group=="BR")
dataPop$group<-as.factor(as.character(dataPop$group))
dataPop$colony<-as.factor(dataPop$colony)
```

## Run wilcox tests comparing both populations
```{r}
willies<-list()
willies[["ML"]]<-wilcox_test(dataPop$ML~dataPop$group,distribution=approximate(nresample=9999))
willies[["SP"]]<-wilcox_test(dataPop$SP~dataPop$group,distribution=approximate(nresample=9999))
willies[["EY"]]<-wilcox_test(dataPop$EY~dataPop$group,distribution=approximate(nresample=9999))
willies[["HL"]]<-wilcox_test(dataPop$HL~dataPop$group,distribution=approximate(nresample=9999))
willies[["CL"]]<-wilcox_test(dataPop$CL~dataPop$group,distribution=approximate(nresample=9999))
```

## Create table object for pops with relevant info
```{r}

datP2<-split(dataPop,f=dataPop$group)
means<-list()
sds<-list()
medians<-list()
mins<-list()
maxs<-list()
cvs<-list()
vars<-list()


for (p in c("BR","TN")){
  for (q in c('ML','SP','EY','HL','CL')){
    means[[p]][q]<-mean(datP2[[p]][,q],na.rm=T)
    sds[[p]][q]<-sd(datP2[[p]][,q],na.rm=T)
    medians[[p]][q]<-median(datP2[[p]][,q],na.rm=T)
    mins[[p]][q]<-min(datP2[[p]][,q],na.rm=T)
    maxs[[p]][q]<-max(datP2[[p]][,q],na.rm=T)
    cvs[[p]][q]<-cv(datP2[[p]][,q])
    vars[[p]][q]<-var(datP2[[p]][,q],na.rm=T)
    
  }
}
m1<-do.call(cbind,means)
colnames(m1)<-c("BR_mean","TN_mean")
m2<-do.call(cbind,sds)
colnames(m2)<-c("BR_sd","TN_sd")
m3<-do.call(cbind,medians)
colnames(m3)<-c("BR_median","TN_median")
m4<-do.call(cbind,mins)
colnames(m4)<-c("BR_min","TN_min")
m5<-do.call(cbind,maxs)
colnames(m5)<-c("BR_max","TN_max")
m6<-do.call(cbind,cvs)
colnames(m6)<-c("BR_cv","TN_cv")
m7<-do.call(cbind,vars)
colnames(m7)<-c("BR_var","TN_var")
all<-cbind(m1,m2,m3,m4,m5,m6,m7)


pvalue<-vector()
mslrStats<-vector()
mslrP<-vector()
wilcoxP<-vector()
wilcoxStats<-vector()
for (q in c('ML','SP','EY','HL','CL')){
  wilcoxP[q]<-pvalue(willies[[q]])[1]
  wilcoxStats[q]<-statistic(willies[[q]])
}

wilcoxStats<-data.frame(wilcoxStats)
wilocxP<-data.frame(wilcoxP)

all<-merge(all,wilcoxP,by="row.names")
row.names(all)<-all$Row.names
allPop<-merge(all[,-1],wilcoxStats,by="row.names")

colnames(allPop)[1]<-"measurement"
colnames(allPop)[colnames(allPop)=="y"]<-"wilcox_p"
write.table(allPop,"/Users/lukas/sciebo/Projects/BSc/Miles Winter/paper/paper/lukas/Figures/Fig2/morphometryTable.populations.tsv",sep="\t",quote=F,row.names = F)

allPop

```

```{r}
## Add significance symbols 
ann_text2  <- data.frame(group = 1.5,SP = Inf,exp = factor("nat. pop.",levels = c("17-DMAG","nat. pop.")))


p2<-p2+  
  geom_text(data = ann_text2,label = "+",vjust=1,size=2.5) + 
  geom_text(data = ann_text2,label = "____",vjust=.7,size=4,fontface = "bold")

ann_text  <- data.frame(group = 1.5,EY = Inf,exp = factor("17-DMAG",levels = c("17-DMAG","nat. pop.")))
ann_text2 <- data.frame(group = 1.5,EY = Inf,exp = factor("nat. pop.",levels = c("17-DMAG","nat. pop.")))

p3<-p3+  
  geom_text(data = ann_text,label = "ø",vjust=1,size=2.5) + 
  geom_text(data = ann_text,label = "____",vjust=.7,size=4,fontface = "bold") +
  geom_text(data = ann_text2,label = "+",vjust=1,size=2.5) + 
  geom_text(data = ann_text2,label = "____",vjust=.7,size=4,fontface = "bold")
  
ann_text  <- data.frame(group = 1.5,HL = Inf,exp = factor("17-DMAG",levels = c("17-DMAG","nat. pop.")))
ann_text2 <- data.frame(group = 1.5,HL = Inf,exp = factor("nat. pop.",levels = c("17-DMAG","nat. pop.")))

p4<-p4+  
  geom_text(data = ann_text,label = "ø",vjust=1,size=2.5) + 
  geom_text(data = ann_text,label = "____",vjust=.7,size=4,fontface = "bold") +
  geom_text(data = ann_text2,label = "+",vjust=1,size=2.5) + 
  geom_text(data = ann_text2,label = "____",vjust=.7,size=4,fontface = "bold")


```

## Plot boxplots for all 4 groups
```{r warning=FALSE}
plot_grid(p1,p2,p3,p4,p5,ncol=3)
dev.print(pdf,"/Users/lukas/sciebo/Projects/BSc/Miles Winter/paper/paper/lukas/Figures/Fig2/Fig2.pdf",width=5,height=4)
```